\bf The Hyperedge Event Model

### 2018-05-16 paper/paper.tex
paper structure change and writing update

\if0\blind
{
  \title{\bf A Hyper-edge Event Model}
  \author{Bomin Kim\thanks{
    The authors gratefully acknowledge \textit{please remember to list all relevant funding sources in the unblinded version}}\hspace{.2cm}\\
    Department of Statistics, Pennsylvania State University,\\
  Aaron Schein\\
  College of Information and Computer Sciences, UMass Amherst,\\
    Bruce Desmarais\\
    Department of Political Science, Pennsylvania State University,\\
    and\\
    Hanna Wallach\\
    Microsoft Research NYC}
  \maketitle
} \fi
\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf A Hyper-edge Event Model}
\end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
We introduce the hyper-edge event model (HeEM)---a generative model for directed edges with multiple recipients or multiple senders. To define the model, we
integrate a dynamic version of the exponential random graph model (ERGM) and generalized linear model (GLM) approach to jointly understand who communicates with whom, and when. We use the model to analyze emails sent between department managers in Montgomery county government in North Carolina. Our application demonstrates that the model is effective at predicting and explaining time-stamped network data involving edges with multiple recipients.
	\end{abstract}

@@

\newpage
\spacingset{1.45} % DON'T change the spacing!
\section{Introduction}\label{sec:intro}
In recent decades, real-time digitized textual communication has developed into a ubiquitous form of social and professional interaction \citep{kanungo2008modeling, szostek2011dealing, burgess2004email, pew2016}. From the perspective of the computational social scientist, this has led to a growing need for methods of modeling interactions that manifest as text exchanged in continuous time. A number of models that build upon topic modeling through Latent Dirichlet Allocation \citep{Blei2003} to incorporate link data as well as textual content have been developed recently \citep{mccallum2005author,lim2013twitter,Krafft2012}. These models are innovative in their extensions that incorporate network information. However, none of the models that are currently available in the literature integrate the rich random-graph structure offered by state of the art models for network structure---such as the exponential random graph model (ERGM) \citep{robins2007introduction,chatterjee2013estimating,hunter2008ergm}. The ERGM is the canonical model for modeling the structure of a static network. It is flexible enough to specify a generative model that accounts for nearly any pattern of edge formation such as reciprocity, clustering, popularity effects \citep{desmarais2017statistical}.

Several network models have been developed that handle time-stamped events in which edge formation is governed by structural dynamics similar to those used in the ERGM \citep{Butts2008,Vu2011,snijders1996stochastic}. Those models are very useful to understand which traits and behaviours are predictive of interactions, however, none of the models explicitly allows hyper-edges---a connection between two or more vertices generated simultaneously\footnote{A hyper-edge connecting just two nodes is simply a usual dyadic edge.}---which is a common property of digitalized textual interactions such as emails and online messages. For instance, \cite{PerryWolfe2012} treat multicast interactions---one type of directed hyper-edge which involve one sender and multiple receivers---via duplication (i.e., obtain pairwise interactions from the original multicast), but their model parameter estimation relies on likelihood approximation which leads to bias. Similarly, \cite{fan2009learning} treats multicast emails as multiple single-recipient edges and randomly jitter the sent times, to avoid the violation of continous-time model assumption. Concentrating on better handling of hyper-edges, we develop the hyper-edge event model (HeEM) which simultaneously models the two components that govern time-stamped event formation: 1) the recipient selection processs that allows multiple senders or receivers, and 2) the time-to-next interactions with flexible distributional choices.

In what follows, we introduce the HeEM by describing how we assume the generatve process of a time-stamped event data (Section 2), and deriving the sampling equations for Bayesian inference (Section 3). Then, we apply the model to the Montgomery county government email data and perform two model validation tasks (Section 4). Finally, the paper finishes in Section 5 with conclusion and discussion.

\section{A Hyper-Edge Event Model (HeEM)}\label{sec:generative process}
Data generated under the model consists of $D$ unique edges. A single edge, indexed by $d \in [D]$, is represented by the three components: the sender $a_d \in [A]$, an indicator vector of recipients $\boldsymbol{r}_d = \{u_{dr} \}_{r=1}^{A}$, and the timestamp $t_d \in (0, \infty)$. For simplicity, we assume that edges are ordered by time such that $t_d \leq t_{d+1}$. While the model can be applied for two type of hyper-edges---edges with (1) a single sender and multiple receivers, and (2) multiple senders and a single receiver---here we only present the generative process for those involving a single sender and multiple receivers (i.e., multicast). For the latter case of hyper-edges, we treat $a_d$ to be an indicator vector of senders $\boldsymbol{a}_d = \{u_{dr} \}_{a=1}^{A}$ and $r_d$ to be the single recipient.
\subsection{Edge Generating Process}\label{subsec: Tie}
For every possible author--recipient pair $(a,r)_{a \neq r}$, we define the ``recipient intensity", which is the likelihood of edge $d$ being sent from $a$ to $r$:
\begin{equation}
\lambda_{adr} = {\boldsymbol{b}}^{\top}\boldsymbol{x}_{adr},
\end{equation}
where $\boldsymbol{b}$ is $P$--dimensional vector of coefficients and $\boldsymbol{x}_{adr}$ is a set of network features which vary depending on the hypotheses regarding canonical processes relevant to network theory such as popularity, reciprocity, and transitivity. In addition, we include intercept term to account for the average (or baseline) number of recipients. We place a Normal prior $\boldsymbol{b} \sim N(\boldsymbol{\mu}_b, \Sigma_b)$.

Next, we hypothesize ``If $a$ were the sender of edge $d$, who would be the recipient(s)?" To do this, we draw each sender's set of recipients from the multivariate Bernoulli (MB) distribution \citep{dai2013multivariate}---a model to estimate the structure of graphs with binary nodes---with probability of 1 being logit($\lambda_{adr}$). In order to avoid the model degeneracy from having an empty recipient set, we define a probability measure ``MB$_{Gibbs}$" motivated by the non-empty Gibbs measure \citep{fellows2017removing} which excludes the all-zero vector from the support of MB distirbution. As a result, we 1) allow multiple recipients or ``multicast", 2) prevent from obtaining zero recipient, and 3) ensure tractable normalizing constant. To be specific, we draw a binary vector $\boldsymbol{u}_{ad}= (u_{ad1},
\ldots, u_{adA})$ 
\begin{equation} \boldsymbol{u}_{ad}  \sim
\mbox{MB}_{Gibbs}(\boldsymbol{\lambda}_{ad}),
\end{equation}
where $\boldsymbol{\lambda}_{id}= \{\lambda_{adr}\}_{r=1}^A$. In particular, we define $\mbox{MB}_{Gibbs}(\boldsymbol{\lambda}_{ad})$ as
\begin{equation}
\begin{aligned}
&p(\boldsymbol{u}_{ad}|\delta, \boldsymbol{\lambda}_{ad}) = \frac{\exp\Big\{\mbox{log}\big(\text{I}( \lVert \boldsymbol{u}_{ad}\rVert_1 > 0 )\big) + \sum_{r\neq a} \lambda_{adr}u_{adr}\Big\}}{Z(\boldsymbol{\lambda}_{ad})} ,
\end{aligned}
\label{eqn:Gibbs}
\end{equation}
where $Z(\boldsymbol{\lambda}_{ad})= \prod_{r \neq a} \big(\mbox{exp}(\lambda_{adr}) + 1\big)-1$ is the normalizing constant and $\lVert \cdot \rVert_1$ is the $l_1$--norm. Again, this is equivalent to assume independent Bernoulli trial on each $u_{adr}$ with probability of 1 being logit($\lambda_{adr}$), excluding the case when all $u_{adr}=0$. We provide the derivation of the normalizing constant as a tractable form in the supplementary material.

@@

In modeling ``when", we do not directly model the timestamp $t_d$. Instead, we assume that each sender's the time-increment or ``time to next interaction" (i.e., $\tau_{d} = t_d-t_{d-1}$) is drawn from a specific distribution in the exponential family.  We follow the generalized linear model (GLM) framework \citep{nelder1972generalized}:
\begin{equation}
\begin{aligned}
E(\tau_{ad}) &= \mu_{ad},\\
V(\tau_{ad}) &= V(\mu_{ad}),
\end{aligned}
\end{equation}
where $\tau_{ad}$ here is a positive real number. Possible choices of distribution include Exponential, Weibull, Gamma, and lognormal\footnote{lognormal distribution is not exponential family but can be used via modeling of $\log(\tau_d)$.} distributions, which are commonly used in time-to-event modeling \citep{rao2000applied,rizopoulos2012joint}. Based on the choice of distribution, we may introduce any additional parameter (e.g., $\sigma_\tau^2$) to account for the variance. We use $f_\tau(\cdot; \mu, \sigma_\tau^2)$ and $F_\tau(\cdot; \mu, \sigma_\tau^2)$ to denote the probability density function (p.d.f) and cumulative density function (c.d.f), respectively, with mean $\mu$ and variance $\sigma^2$.

@@

P(\boldsymbol{\eta}, \sigma_\tau^2| \boldsymbol{u}, \boldsymbol{y})\propto \prod_{m=1}^M\Big(\prod_{d:t_d=t_m^*}f_{\tau}(t_m^*-t_{m-1}^*; \mu_{a_d d}, \sigma_\tau^2)\times \prod_{a \notin \{a_d\}_{d:t_d=t_m^*}}\big(1-F_{\tau}(t_m^*-t_{m-1}^*; \mu_{a d}, \sigma_\tau^2) \big)\Big),

@@

 				update $\boldsymbol{\eta}$ using Metropolis-Hastings --- Equation (10)

@@

 	   Our data come from the North Carolina county government email dataset collected by \cite{ben2017transparency} that includes internal email corpora covering the inboxes and outboxes of managerial-level employees of North Carolina county governments. Out of over twenty counties, we chose Montgomery County to 1) test our model using data with a lot of hyper-edges (16.76\%), all of which are a single sender multiple receiver cases, and 2) limit the scope of this initial application. To summarize, Montgomery County email network contains 680 emails, sent and received by 18 department managers over a period of 3 months (March--May) in 2012. 

@@

	In the example of email networks, we form the covariate vector $\boldsymbol{x}_{adr}$ using time-varying network statistics based on the time interval prior to and including $t_{d-1}$. Along with intercept term, we compute the network statistics \citep{PerryWolfe2012}, where the time interval tracks 7 days prior to the last email was sent $l_d = (t_{d-1}-7\mbox{days}, t_{d-1}]$. To be specific, we use

@@

		\item[1.] intercept ${x}_{adr1} =1$;
		\item[2.] outdegree ${x}_{adr2} =\sum\limits_{d^\prime: t_{d^\prime} \in l_d} I(a_{d^\prime} = a)$;
		\item[3.] indegree ${x}_{adr3}=\sum\limits_{d^\prime: t_{d^\prime} \in l_d} I(u_{d^\prime r} = 1)$;
		\item[4.] send ${x}_{adr4}=\sum\limits_{d^\prime: t_{d^\prime} \in l_d} I(a_{d^\prime} = a)I(u_{d^\prime r} = 1)$;
		\item[5.] receive ${x}_{adr5}=\mbox{send}(r,a)$;
		\item[6.] 2-send ${x}_{adr6} = \sum\limits_{h \neq a, r} \mbox{send}(a,h)\mbox{send}(h,r)$;
		\item[7.] 2-receive ${x}_{adr7}= \sum\limits_{h \neq a, r} \mbox{send}(h,a)\mbox{send}(r,h)$;
		\item[8.] sibling ${x}_{adr8}=\sum\limits_{h \neq a, r} \mbox{send}(h,a)\mbox{send}(h,r)$;
		\item[9.] cosibling ${x}_{adr9}=\sum\limits_{h \neq a, r} \mbox{send}(a,h)\mbox{send}(r,h)$;
		\item[10.] Nreceive ${x}_{adr10}=\sum\limits_{d^\prime: t_{d^\prime} \in l_d} \sum\limits_{r=1}^A I(a_{d^\prime} = a)I(u_{d^\prime r} = 1)$;
		\item[11.] outdegree$*$Nreceive ${x}_{adr11} = {x}_{adr2}\times{x}_{adr10};$

@@

	where $I(\cdot)$ is an indicator function, and the last covariate is the interaction term between outdegree and Nreceive. As with the nodal (2, 3, 10, 11) and dyadic (4, 5) effects, the triadic (6--9) effects are designed so that their coefficient have a straightforward interpretation, which are illustrated in Figure \ref{figure:netstats}.
		\begin{figure}[H]

@@

For time-related covariates $\boldsymbol{y}_{ad}$ in Section \ref{subsec:Time}, we form the covariate vector $\boldsymbol{y}_{ad}$ using the seven statistics which may possibly effect ``time to send the next document". Note that these statistics could depend on $a$ only (nodal effect), $d$ (or $t_d$) only (temporal effect), or both $a$ and $d$ (nodal \& temporal effect). Specifically, the statistics are defined as
\begin{itemize}
	\spacingset{1}
	\item[1.] intercept ${y}_{ad1} =1$;
	\item[2.] outdegree ${y}_{ad2} =\sum\limits_{d^\prime: t_{d^\prime} \in l_d} I(a_{d^\prime} = a)$;
	\item[3.] indegree ${y}_{ad3}=\sum\limits_{d^\prime: t_{d^\prime} \in l_d} I(u_{d^\prime a} = 1)$;
	\item[4.] gender ${y}_{ad4}=I(a= \mbox{female})$;
	\item[5.] manager ${y}_{ad5}=I(a= \mbox{County Manager})$;
	\item[6.] weekend ${y}_{ad6} = I(t_{d-1}= \mbox{weekend})$;
	\item[7.] am/pm ${y}_{ad7}= I(t_{d-1}= \mbox{pm})$.
\end{itemize}
\subsection{Parameter Estimation and Interpretation}\label{subsec:Result_email}
Based on some preliminary experiments, we choose lognormal distribution to model time-to-next-email in Montgomery county email data, thus we estimate the variance parameter $\sigma_\tau^2$ with the prior specified as inverse-Gamma($2, 1$). In this section, we present the results based on 55,000 MCMC outer iterations with a burn-in of 15,000, where we thin by keeping every 40th sample. While the inner iterations for $\boldsymbol{\eta}$ and $\sigma_\tau^2$ are fixed as 1, we use 10 inner iterations for $\boldsymbol{b}$ to adjust for its slower convergence rate.
	\begin{figure}[!b]
	\centering
	\includegraphics[width=1\textwidth]{plots_paper/best-1.png}	
	\caption {Posterior distribution of $\boldsymbol{b}$ estimates.}
	\label{figure:betaresults}
\end{figure}

Figure \ref{figure:betaresults} shows the boxplots summarizing posterior samples of $\boldsymbol{b}$. Since we use logit probability of $\lambda_{adr}$ for edge generating process (Section \ref{subsec: Tie}), we have
\begin{equation*}
\mbox{logit}(\lambda_{adr})=\log\Big(\frac{\lambda_{adr}}{1-\lambda_{adr}}\Big) =b_{1}+b_{2} x_{adr2}\ldots+b_{11}x_{adr11},
\end{equation*}
and can interpret the parameter estimates in terms of odds ratio $\frac{\lambda_{adr}}{1-\lambda_{adr}}=\exp(b_{1}+b_{2} x_{adr2}\ldots+b_{11}x_{adr11})$.
Firstly, we can see that the effect of send (i.e., ``number of times the actor $a$ sent emails to the actor $r$ over the last week'') is positive, implying that if the actor $a$ sent $n$ number of emails to $r$ last week, then the actor $a$ is approximately $\exp(0.274\times n)\approx(1.315)^n$ times more likely to send an email to $r$. Similar interpretation can be applied for the statistics indegree and Nrecieve. For example, as the actor $r$ received $n$ emails and the actor $a$ sent emails to $n$ rececievers over the last week, the actor $a$ is $\exp(0.086\times n)\approx(1.091)^n $ times and $\exp(0.047\times n)\approx(1.048)^n$ times, respectively, more likely to send an email to $r$. On the other hand, outdegree statistic has negative effect---i.e., if the actor $a$ sent $n$ number of emails to anyone last week, then the actor $a$ is approximately $\exp(-0.109\times n)\approx(0.897)^n$ times likely to send an email to $r$---possibly due to its multicollinearity with the statistics send and Nreceive. This can be interpreted as ``if the actor $a$ sent large number of emails over the last week but 1) not to $r$ (excluding the effect of send) or 2) no broadcast or multicast emails (excluding the effect of Nreceive)'' then the actor $a$ is less likely to send an email to $r$. Rest of statistics are not statistically significant because their 95\% credible interval do not include 0. Surprisingly, none of the triadic effects seem to have significant effect on the edge generating process.

>>>>

\if0\blind
{
  \title{\bf The Hyperedge Event Model}
  \author{Bomin Kim\thanks{
    The authors gratefully acknowledge \textit{please remember to list all relevant funding sources in the unblinded version}}\hspace{.2cm}\\
    Department of Statistics, Pennsylvania State University,\\
  Aaron Schein\\
  College of Information and Computer Sciences, UMass Amherst,\\
    Bruce Desmarais\\
    Department of Political Science, Pennsylvania State University,\\
    and\\
    Hanna Wallach\\
    Microsoft Research NYC}
  \maketitle
} \fi
\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf The Hyperedge Event Model}
\end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
We introduce the hyperedge event model (HEM)---a generative model for directed edges with multiple recipients or multiple senders. To define the model, we
integrate a dynamic version of the exponential random graph model (ERGM) and generalized linear model (GLM) approach to jointly understand who communicates with whom, and when. We use the model to analyze emails sent between department managers in Montgomery county government in North Carolina. Our application demonstrates that the model is effective at predicting and explaining time-stamped network data involving edges with multiple recipients.
	\end{abstract}

@@

\newpage
\spacingset{1.45} % DON'T change the spacing!
\section{Introduction}\label{sec:intro}
In recent decades, real-time digitized textual communication has developed into a ubiquitous form of social and professional interaction \citep{kanungo2008modeling, szostek2011dealing, burgess2004email, pew2016}. From the perspective of the computational social scientist, this has led to a growing need for methods of modeling interactions that manifest as edges exchanged in continuous time. A number of models that build upon topic modeling through Latent Dirichlet Allocation \citep{Blei2003} to incorporate link data as well as textual content have been developed recently \citep{mccallum2005author,lim2013twitter,Krafft2012}. These models are innovative in their extensions that incorporate network information. However, none of the models that are currently available in the literature integrate the rich random-graph structure offered by state of the art models for network structure---such as the exponential random graph model (ERGM) \citep{robins2007introduction,chatterjee2013estimating,hunter2008ergm}. The ERGM is the canonical model for modeling the structure of a static network. It is flexible enough to specify a generative model that accounts for nearly any pattern of edge formation such as reciprocity, clustering, popularity effects \citep{desmarais2017statistical}.

Several network models have been developed that handle time-stamped events in which edge formation is governed by structural dynamics similar to those used in the ERGM \citep{Butts2008,Vu2011,snijders1996stochastic}. Those models are very useful to understand which traits and behaviours are predictive of interactions, however, none of the models explicitly allows hyperedges---a connection between two or more vertices generated simultaneously\footnote{A hyperedge connecting just two nodes is simply a usual dyadic edge.}---which is a common property of digitalized textual interactions such as emails and online messages. For instance, \cite{PerryWolfe2012} treat multicast interactions---one type of directed hyperedge which involve one sender and multiple receivers---via duplication (i.e., obtain pairwise interactions from the original multicast), but their model parameter estimation relies on likelihood approximation which leads to bias. Similarly, \cite{fan2009learning} treats multicast emails as multiple single-recipient edges and randomly jitter the sent times, to avoid the violation of continous-time model assumption. On the other hand, there have been numerous studies for hyperedges under the hypergraph theory \citep{karypis1999multilevel} in phyics literature, including a random hypergraph model \citep{ghoshal2009random}, but most of them are designed specifically for engineering applications---e.g., multiple tagging networks \citep{zlatic2009hypergraph,zhang2010hypergraph}, cellular networks \citep{klamt2009hypergraphs}, and personalized recommendations \citep{zhang2010personalized,blattner2009b}. To our knowledge, there is a lack of extensive efforts to incorportate hyperedge modeling into general statistical framework.

Concentrating on developing a statistical network model with better handling of hyperedges, we introduce the hyperedge event model (HEM) which simultaneously models the two components that govern time-stamped event formation: 1) the recipient selection processs that allows multiple senders or receivers, and 2) the time-to-next interactions with flexible distributional choices. In what follows, we introduce the HEM by describing how we assume the generatve process of a time-stamped event data (Section 2), and deriving the sampling equations for Bayesian inference (Section 3). Then, we apply the model to the Montgomery county government email data and perform two model validation tasks (Section 4). Finally, the paper finishes in Section 5 with conclusion and discussion.

\section{The Hyperedge Event Model (HEM)}\label{sec:generative process}
Data generated under the model consists of $D$ unique edges. A single edge, indexed by $d \in [D]$, is represented by the three components: the sender $a_d \in [A]$, an indicator vector of recipients $\boldsymbol{r}_d = \{u_{dr} \}_{r=1}^{A}$, and the timestamp $t_d \in (0, \infty)$. For simplicity, we assume that edges are ordered by time such that $t_d \leq t_{d+1}$. While the model can be applied for two type of hyperedges---edges with (1) a single sender and multiple receivers, and (2) multiple senders and a single receiver---here we only present the generative process for those involving a single sender and multiple receivers (i.e., multicast). For the latter case of hyperedges, we treat $a_d$ to be an indicator vector of senders $\boldsymbol{a}_d = \{u_{dr} \}_{a=1}^{A}$ and $r_d$ to be the single recipient.
\subsection{Edge Generating Process}\label{subsec: Tie}
For every possible author--recipient pair $(a,r)_{a \neq r}$, we define the ``recipient intensity", which is the likelihood of edge $d$ being sent from $a$ to $r$:
\begin{equation}
\lambda_{adr} = {\boldsymbol{b}}^{\top}\boldsymbol{x}_{adr},
\end{equation}
where $\boldsymbol{b}$ is $P$--dimensional vector of coefficients and $\boldsymbol{x}_{adr}$ is a set of network features which vary depending on the hypotheses regarding canonical processes relevant to network theory such as popularity, reciprocity, and transitivity. In addition, we include intercept term to account for the average (or baseline) number of recipients. We place a Normal prior $\boldsymbol{b} \sim N(\boldsymbol{\mu}_b, \Sigma_b)$.

Next, we hypothesize ``If $a$ were the sender of edge $d$, who would be the recipient(s)?" To do this, we draw each sender's set of recipients from the multivariate Bernoulli (MB) distribution \citep{dai2013multivariate}---a model to estimate the structure of graphs with binary nodes---with probability of 1 being logit($\lambda_{adr}$). In order to avoid the model degeneracy from having an empty recipient set, we define a probability measure ``MB$_{G}$" motivated by the non-empty Gibbs measure \citep{fellows2017removing} which excludes the all-zero vector from the support of the multivariabe Bernoulli distirbution. As a result, we 1) allow multiple recipients or ``multicast", 2) prevent from obtaining zero recipient, and 3) ensure tractable normalizing constant. To be specific, we draw a binary vector $\boldsymbol{u}_{ad}= (u_{ad1},
\ldots, u_{adA})$ 
\begin{equation} \boldsymbol{u}_{ad}  \sim
\mbox{MB}_{G}(\boldsymbol{\lambda}_{ad}),
\end{equation}
where $\boldsymbol{\lambda}_{id}= \{\lambda_{adr}\}_{r=1}^A$. In particular, we define $\mbox{MB}_{G}(\boldsymbol{\lambda}_{ad})$ as
\begin{equation}
\begin{aligned}
&p(\boldsymbol{u}_{ad}|\delta, \boldsymbol{\lambda}_{ad}) = \frac{\exp\Big\{\mbox{log}\big(\text{I}( \lVert \boldsymbol{u}_{ad}\rVert_1 > 0 )\big) + \sum_{r\neq a} \lambda_{adr}u_{adr}\Big\}}{Z(\boldsymbol{\lambda}_{ad})} ,
\end{aligned}
\label{eqn:Gibbs}
\end{equation}
where $Z(\boldsymbol{\lambda}_{ad})= \prod_{r \neq a} \big(\mbox{exp}(\lambda_{adr}) + 1\big)-1$ is the normalizing constant and $\lVert \cdot \rVert_1$ is the $l_1$--norm. Again, this is equivalent to assume independent Bernoulli trial on each $u_{adr}$ with probability of 1 being logit($\lambda_{adr}$), excluding the case when all $u_{adr}=0$. We provide the derivation of the normalizing constant as a tractable form in the supplementary material.

@@

In modeling ``when", we do not directly model the timestamp $t_d$. Instead, we assume that each sender's the time-increment or ``time to next interaction" (i.e., $\tau_{d} = t_d-t_{d-1}$) is drawn from a specific distribution in the exponential family.  We follow the generalized linear model (GLM) framework \citep{nelder1972generalized}:
\begin{equation}
\begin{aligned}
E(\tau_{ad}) &= \mu_{ad},\\
V(\tau_{ad}) &= V(\mu_{ad}),
\end{aligned}
\end{equation}
where $\tau_{ad}$ here is a positive real number. Possible choices of distribution include exponential, Weibull, gamma, and lognormal\footnote{lognormal distribution is not exponential family but can be used via modeling of $\log(\tau_d)$.} distributions, which are commonly used in time-to-event modeling \citep{rao2000applied,rizopoulos2012joint}. Based on the choice of distribution, we may introduce any additional parameter (e.g., $\sigma_\tau^2$) to account for the variance. We use $f_\tau(\cdot; \mu, \sigma_\tau^2)$ and $F_\tau(\cdot; \mu, \sigma_\tau^2)$ to denote the probability density function (p.d.f) and cumulative density function (c.d.f), respectively, with mean $\mu$ and variance $\sigma^2$.

@@

\begin{aligned}
&P(\boldsymbol{\eta}, \sigma_\tau^2| \boldsymbol{u}, \boldsymbol{y})\\&\propto \prod_{m=1}^M\Big(\prod_{d:t_d=t_m^*}f_{\tau}(t_m^*-t_{m-1}^*; \mu_{a_d d}, \sigma_\tau^2)\times \prod_{a \notin \{a_d\}_{d:t_d=t_m^*}}\big(1-F_{\tau}(t_m^*-t_{m-1}^*; \mu_{a d}, \sigma_\tau^2) \big)\Big),
\end{aligned}

@@

 				update $\boldsymbol{\eta}$ using Metropolis-Hastings --- Equation (10)\\

@@

 	   Our data come from the North Carolina county government email dataset collected by \cite{ben2017transparency} that includes internal email corpora covering the inboxes and outboxes of managerial-level employees of North Carolina county governments. Out of over twenty counties, we chose Montgomery County to 1) test our model using data with a lot of hyperedges (16.76\%), all of which are a single sender multiple receiver cases, and 2) limit the scope of this initial application. To summarize, Montgomery County email network contains 680 emails, sent and received by 18 department managers over a period of 3 months (March--May) in 2012. 

@@

	In the example of email networks, we form the covariate vector $\boldsymbol{x}_{adr}$ using time-varying network statistics based on the time interval prior to and including $t_{d-1}$. Given a set of interaction data, a primary modeling goal lies in determining which characteristics and behaviors of the senders and receivers are predictive of interaction. This email application specifically give rise to the following question: ``To what extent are nodal, dyadic or triadic network effects, as characterized
by past interaction behaviors, relevant to predicting future events?" We employ the network statistics in \cite{PerryWolfe2012}, along with the intercept term, to address the question. Specifically, our time interval tracks 7 days prior to the last email was sent $l_d = (t_{d-1}-7\mbox{ days}, t_{d-1}]$. For $a \in [A], r \in [A]$, and $d \in [D]$, we define the effects

@@

		\item[1.] intercept: ${x}_{adr1} =1$;
		\item[2.] outdegree of sender $a$: ${x}_{adr2} =\sum_{d^\prime: t_{d^\prime} \in l_d} I(a_{d^\prime} = a)$;
		\item[3.] indegree of receiver $r$: ${x}_{adr3}=\sum_{d^\prime: t_{d^\prime} \in l_d} I(u_{d^\prime r} = 1)$;
		\item[4.] send: ${x}_{adr4}=\sum_{d^\prime: t_{d^\prime} \in l_d} I(a_{d^\prime} = a)I(u_{d^\prime r} = 1)$;
		\item[5.] receive: ${x}_{adr5}=\mbox{send}(r,a)$;
		\item[6.] 2-send: ${x}_{adr6} = \sum_{h \neq a, r} \mbox{send}(a,h)\mbox{send}(h,r)$;
		\item[7.] 2-receive: ${x}_{adr7}= \sum_{h \neq a, r} \mbox{send}(h,a)\mbox{send}(r,h)$;
		\item[8.] sibling: ${x}_{adr8}=\sum_{h \neq a, r} \mbox{send}(h,a)\mbox{send}(h,r)$;
		\item[9.] cosibling: ${x}_{adr9}=\sum_{h \neq a, r} \mbox{send}(a,h)\mbox{send}(r,h)$;
		\item[10.] hyperedge size of sender $a$: ${x}_{adr10}=\sum_{d^\prime: t_{d^\prime} \in l_d} \sum_{r=1}^A I(a_{d^\prime} = a)I(u_{d^\prime r} = 1)$;
		\item[11.] outdegree$*$hyperedge size: ${x}_{adr11} = {x}_{adr2}\times{x}_{adr10};$

@@

	where $I(\cdot)$ is an indicator function, and the last covariate is the interaction term between outdegree and hyperedge size. As with the nodal (2, 3, 10, 11) and dyadic (4, 5) effects, the triadic (6--9) effects are designed so that their coefficient have a straightforward interpretation. Dyadic statistics ``send" and ``receive" are defined as above such that these covariates measure the number of emails sent from $a$ to $r$ and $r$ to $a$, respectively, within the last 7 days. In the example of triadic statistics, the covariate ``2-send" counts the pairs of emails involving some node $h$ distinct from $a$ and $r$ such that an email from $a$ to $h$ and $h$ to $r$ are separately observed within the last 7 days. Other triadic covariates behave similarly, and their interpretations are also analogous, which are illustrated in Figure \ref{figure:netstats}.
		\begin{figure}[!t]

@@

For time-related covariates $\boldsymbol{y}_{ad}$ in Section \ref{subsec:Time}, we form the covariate vector $\boldsymbol{y}_{ad}$ using the seven statistics which may possibly effect ``time to send the next document". Note that these statistics could depend on $a$ only (nodal effect), $d$ (or $t_d$) only (temporal effect), or both $a$ and $d$ (nodal \& temporal effect). Specifically, the statistics are defined as
\begin{itemize}
	\spacingset{1}
	\item[1.] intercept: ${y}_{ad1} =1$;
	\item[2.] outdegree of sender $a$: ${y}_{ad2} =\sum_{d^\prime: t_{d^\prime} \in l_d} I(a_{d^\prime} = a)$;
	\item[3.] indegree of sender $a$: ${y}_{ad3}=\sum_{d^\prime: t_{d^\prime} \in l_d} I(u_{d^\prime a} = 1)$;
	\item[4.] gender of sender $a$: ${y}_{ad4}=I(a= \mbox{female})$;
	\item[5.] manager status of sender $a$: ${y}_{ad5}=I(a= \mbox{County Manager})$;
	\item[6.] weekend of event $d$: ${y}_{ad6} = I(t_{d-1}= \mbox{weekend})$;
	\item[7.] am/pm of event $d$: ${y}_{ad7}= I(t_{d-1}= \mbox{pm})$.
\end{itemize}

	\subsection{Prediction Experiment}\label{subsec:Experiment_email}

	The HEM has many component parts that need to be specified by the user (i.e., the selection of the event timing features $\bm{y}$, the recipient selection features $\bm{x}$, and the event time distribution $f$). Many of these components will be specified based on user expertise (e.g., regarding which features would drive recipient selection), but some decisions may require a data-driven approach to model specification. For example, though theoretical considerations may inform the specification of features, subject-matter expertise is unlikely to inform the decision regarding the family of the event time distribution. Furthermore, since different distribution families (and model specifications more generally) may involve different size parameter spaces, any data-driven approach to model comparison must guard against over-fitting the data. In this section we present a general-purpose approach to evaluating the HEM specification using out-of-sample prediction. We illustrate this approach by comparing alternative distributional families for the event timing component of the model. Here, we specifically compare the predictive performance from two distributions---lognormal and exponential. We particularly choose exponential distribution as an alternative to what we used earlier (i.e., lognormal) since exponential is the most commonly specified distribution for time-to-event data which is also used in the stochastic actor oriented models \citep{snijders1996stochastic} as well as their extensions \citep{snijders2007modeling}.

		We evaluate the model's ability to predict edges and timestamps from the Montgomery County email data, conditioned on their ``training" part of the data. To perform the experiment, we separately formed a test split of each three components---sender, recipients, and timestamps---by randomly selecting ``test" data with probability $p=0.1$. Any missing variables were imputed by drawing samples from their conditional posterior distributions, given the observed data, parameter estimates, and current values of test data. We then run inference to update the latent variables given the imputed and observed data. We iterate the two steps---imputation and inference---multiple times to obtain enough number of estimates for ``test" data. Algorithm \ref{alg:PPE} outlines this procedure in detail. We run the experiment and measure the predictive performance of two separate time distributions using 500 predicted samples, which are depicted in Figure \ref{figure:PPEresults}. First, we compare the correct sender probability for missing documents $\{d:a_d=\mbox{NA}\}$, which corresponds to $\pi_{a_{d}}=P(a_{d} = a^{obs}_{d}|\cdot)$ in Algorithm \ref{alg:PPE}. The plot on the left shows that lognormal distribution reaches higher correct sender probability than exponential distribution, although their centers (e.g., medians) are similar. Secondly, we compute similar measure for recipients---correct recipient probability $P(r_{dr} = r^{obs}_{dr})$ in Algorithm \ref{alg:PPE} where $r^{obs}_{dr}$ is either 1 (received) or 0 (not received)---and visually compare the two distributions in logit scale (i.e., $\log(\frac{P(r_{dr} = r^{obs}_{dr}|\cdot)}{1-P(r_{dr} = r^{obs}_{dr})})$). Since the edge generating process (Section \ref{subsec: Tie}) is not directly affected by the choice of timestamp distribution, there is no significant difference between lognormal and exponential in their performance of predicting unknown recipients. Finally, prediction errors for missing timestamps are measured using the median absolute percentage error (MdAPE) \citep{hyndman2006another}, where we take the median of absolute percentage errors $p_{di}$ for $i=1,\ldots,500$ for $d$th document, calculated as follows:
		\begin{algorithm}[!t]
			\caption{Out-of-Sample Predictions}
			\label{alg:PPE}
			\begin{algorithmic}
				\STATE {\bfseries Input:} data $ \{ (a_d, \boldsymbol{r}_d, t_d)\}_{d=1}^D$, 
				number of new data to generate $R$,
				hyperparameters
				\vskip 0.1in
				\textbf{Test splits:}	
				\STATE draw test authors with $p=0.1$ (out of $D$ authors) 
				\STATE draw test recipients
				with $p=0.1$ (out of $D\times (A-1)$ recipient indicators $\{\{\boldsymbol{r}_{dr}\}_{r\in [A]_{\backslash a_d}}\}_{d=1}^D$)
				\STATE draw test timestamps with $p=0.1$  (out of $D$ timestamps) 
				\STATE set the ``test" data as ``missing" (NA)
				\vskip 0.1in
				\textbf{Imputation and inference:}	
				\STATE initialize the parameters $(\boldsymbol{b}, \boldsymbol{\eta}, \boldsymbol{u}, \sigma^2_\tau)$
				\FOR{$r=1$ {\bfseries to}  $R$}
				\FOR{$d=1$ {\bfseries to}  $D$}
				\IF{$a_d=$  NA}
				\FOR{$a=1$ {\bfseries to} $A$}
				\STATE compute $\pi_{a} $ using $P(a_d= a | \cdot)=f_{\tau}(\tau_{d}; \mu_{a_dd}, \sigma_\tau^2)\times \prod_{a\neq a_d}\big(1-F_{\tau}(\tau_{d}; \mu_{a d}, \sigma_\tau^2) \big)$
				\ENDFOR
